The Werewolf Among Us: Humans vs LLMs in Multi-Agent Games

### 2025-05-05 Quarto/paper.qmd

---
title: La Palma Earthquakes
author:
  - name: Steve Purves
    orcid: 0000-0002-0760-5497
    corresponding: true
    email: steve@curvenote.com
    roles:
      - Investigation
      - Project administration
      - Software
      - Visualization
    affiliations:
      - Curvenote
  - name: Rowan Cockett
    orcid: 0000-0002-7859-8394
    corresponding: false
    roles: []
    affiliations:
      - Curvenote
keywords:
  - La Palma
  - Earthquakes
abstract: |
  In September 2021, a significant jump in seismic activity on the island of La Palma (Canary Islands, Spain) signaled the start of a volcanic crisis that still continues at the time of writing. Earthquake data is continually collected and published by the Instituto Geográphico Nacional (IGN). ...
date: last-modified
citation:
  container-title: Earth and Space Science
---

# Example site and pdf

Lorem ipsum

Test update

<!-- This shouldnt be seen, its a comment -->

## Example Subsection

Moar lorem ipsum

# Another Section

>>>>

---
title: "The Werewolf Among Us: Humans vs LLMs in Multi-Agent Games"
author:
  - name: Bhavana Jonnalagadda
    orcid: 0009-0009-1054-0995
    corresponding: true
    email: bhavana.jonnalagadda@colorado.edu
    affiliations:
      - name: University of Colorado Boulder
        id: 1
        city: Boulder
        state: CO
  - name: Riley Jones
    email: rjones@colorado.edu
    affiliations:
      - id: 1
keywords:
  - social deduction games
  - persuasion modeling
  - Werewolf Among Us dataset
  - large language models
  - multimodal analysis
abstract: |
  Abstract TODO
date: last-modified
citation:
  container-title: Human vs LLM Multi-Agent Games
---

::: {.content-visible when-format="html"}
::: {.column-margin}
![](Images/logo.png)
:::
:::

# Introduction

<!-- TODO: Broken for some reason?!?!? Fix -->
<!-- ::: {.content-hidden unless-format="latex"}
::: {.callout-tip}
## View Online
This PDF's content, the source code notebooks, along with live interactive versions of all figures (and more!) is available at the website version of this paper: [https://cuboulder-ds.github.io/CSCI-5423-Final](https://cuboulder-ds.github.io/CSCI-5423-Final)
:::
::: -->

- A description of the problem and its significance
- How do LLMs function in a multi-agent environment?
  - Each has limited information
  - Approximately the same ability, traits, skills
- We use LLMs to simulate whether synthetic agents can participate in complex, adversarial group dynamics
- Werewolf is a good candiate for testing multi agent systems of cooperation and secrecy
  - Need citation
  - The game tests adaptive reasoning, strategic alignment, and collective threat detection under special conditions

## Related Work

### Multi-Agent LLMs

- Among us game [@chiAMONGAGENTSEvaluatingLarge2024]
- Collective problem solving [@duLargeLanguageModels2024]
  - "analyses indicate that LLM agent groups exhibit more disagreements, complex statements, and a propensity for positive statements compared to human groups"
- Govsim [@piattiCooperateCollapseEmergence2024]
  - "In GOVSIM, a society of AI agents must collectively balance exploiting a common resource with sustaining it for future use. This environment enables the study of how ethical considerations, strategic planning, and negotiation skills impact cooperative outcomes."
- All found similar themes
  - That LLMs are capable and good at understanding the rules
  - That they can cooperate and be sneaky

### LLMs and Werewolf

- Examination of improving werewolf by LLMs [@xuLanguageAgentsReinforcement2024]
  - "our agents use an LLM to perform deductive reasoning and generate a diverse set of action candidates. Then an RL policy trained to optimize the decision-making ability chooses an action from the candidates to play in the game. Extensive experiments show that our agents overcome the intrinsic bias and outperform existing LLM-based agents in the Werewolf game."
- Werewolf Arena [@bailisWerewolfArenaCase2024]
  - Used in this paper
- Explicitly discuss how none of the exisiting LLM+Werewolf papers examine the differences/compare from a human dataset

# Methods

## Data

### Werewolf Among Us Human Dataset

- Human dataset description [@laiWerewolfUsMultimodal2022]
- Is specifically for a form of one-night werewolf
  - Describe key differences
- Used specifically for the text available
  - and annotations of persuasion strategy on the text

### Werewolf Arena

- [@bailisWerewolfArenaCase2024]
- Discuss the framework, how it works, prompts, etc
- Discuss what types of runs we did
- Discuss the data included in output
- Talk about how we had to annotate the LLM speech with persuasion strategies ourselves

## Analysis

- Formatted data to match, performed various comparisons

# Results

{{< embed Data/EDA_Comparison.ipynb#fig-llm-winrounds >}}

{{< embed Data/EDA_Comparison.ipynb#fig-stratbyrole-hum >}}
{{< embed Data/EDA_Comparison.ipynb#fig-stratbyrole-llms >}}

# Discussion and Conclusion

Interpret findings, discuss limitations, and propose future work.

## Limitations

## Future Work

## Summary

Summarize contributions and insights from the project.

---

<!-- See https://quarto.org/docs/authoring/citations.html -->
# References

::: {#refs}
:::

---

# Project Contributions

**Bhavana Jonnalagadda**:

- Paper framework (Quarto) setup
- Github repo management
- EDA on LLM dataset
- Final comparison EDA and results analysis
- Results section
- Discussion and Conclusion section
- Abstract

**Riley Jones**:

- EDA on human dataset
- Werewolf Arena LLM simulation running and data aquisition
- Introduction section
- Methods section
